---
title: "p4-Dilutions-4-rep-combo"
format: gfm
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width=7, fig.height = 9)
```

Load in packages needed for the analysis

```{r}
library("survival")
library("survminer")
library(dplyr)
library(tidyr)
library(stringr)
library(AICcmodavg)
```

### Loop to convert the example data.frame 'df' into properly formatted data.frame 'results'

```{r}
#write a function to transform a data.frame that has the column format 'vial | treatment | D0 | D1 | D2...', with one row for each vial
#into a long version in tidy format that can be input to make a survivorship curve
convert_df<-function(df){
  #open empty data.frame to store results
  results<-data.frame(vial=character(),treatment=character(),dead=numeric(),status=numeric())
  #This loop will index out each row (one row per vial) one at a time, transform it into long format (one row per fly), and add the information to the empty data.frame called results
  for(i in 1:nrow(df)){
  #isolate the row (vial) you want to work on
  temp<-df[i,]
    #iteratively isolate each day for this vial (day 0 must be column 3, day 1 column 4, etc.). Loop stops the column before the last day
    for(j in 3:(ncol(temp)-1)){
      #assign the number of flies that died in the vial on that day (starting with day 1) to the variable 'z'
      z<-temp[1,j]-temp[1,j+1]
        #if >0 flies died add this information to the results dataframe
        if(z>0){
          #iterate over 1 through total number of dead flies
          for(k in 1:z){
            #add a new row to the 'results' data.frame for the given dead fly, specifying vial #, treatment, day died, and
            #record the current vial #
            vial<-temp[,1]
            #record the genotype of the current vial
            treatment<-temp[,2]
            #record the death date of the flies that died on this day (assumes that your input DF starts with day 0 in column 3)
            dd<-j-2
            #append this information into a new row in the 'results' data.frame, and add a '1' in the 4th column to indicate mortality
            results[nrow(results)+1,]<- c(vial,treatment,dd,1)
          } #close for loop
        } #close if loop
    } #close for loop
  
  #now assign the number of flies remaining in the vial on the last day (value in the last column of the row) to the variable 'z'
  z<-temp[1,j+1]
    #if there are any flies alive in the vial on the last day
    if(z>0){
      #iterate over 1:(number of flies alive on the last day)
      for(l in 1:z){
        #record the current vial #
        vial<-temp[,1]
        #record the genotype of the current vial
        treatment<-temp[,2]
        #record the last day we recorded this fly alive (assumes that your input DF starts with day 0 in column 3)
        dd<-j-2
        #append this information into a new row in the 'results' data.frame, and add a '0' in the 4th column to indicate that the fly made it to the end of the experiment
        results[nrow(results)+1,]<- c(vial,treatment,dd,0)
      } #close for loop
    } #close if loop
  } #close original for loop
results$dead<-as.numeric(results$dead)  #reiterate that this column must be class numeric
results$status<-as.numeric(results$status)  #reiterate that this column must be class numeric
results$vial <- as.factor(results$vial) # make sure vial is considered a factor
# gives you only the results dataframe as output from function 
return(results) 
} #close function
```

Read in raw data

**Note that for these datasets, the adjusted N number is the number of flies alive on day 2 because there seemed to be more early death in these experiments than usual**

```{r}
#read the file from csv
df<-read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/Infection_survival_analyses/20240209-dilution-rep-4/dilution-infection-combo-sheet.csv")

# Remove the rows of the undiluted samples because I did not do those for the other replicates
df <- df[c(1:12, 15:56),]
# separate out columns needed
df <- df[,c(1,3,15:29)]

```

Convert dataframe

```{r}
df.convert<-convert_df(df)
```

Plot survivial curve with every line separate, yes this is a mess

```{r}
# change to not have confidence intervals in this one so you can see them 
df_fit<- survfit(Surv(dead, status) ~ treatment, data=df.convert)
ggsurvplot(df_fit,
          pval = FALSE, conf.int = FALSE,
          #risk.table = TRUE, # Add risk table
          #risk.table.col = "strata", # Change risk table color by groups
          #linetype = "strata", # Change line type by groups
          #surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw()) # Change ggplot2 theme
          # palette = c("orchid", "aquamarine", "blueviolet", "darkslategray3")) + ylab("Survival Proporation") + xlab("Days post injection")

```

Add in columns for sex and dilution separately

```{r}
# add in sex information and DiNV information as separate columns by splitting the treatment column 
# split the columns 
df.convert_S <- str_split_fixed(df.convert$treatment, "-", 2)

# change column names
colnames(df.convert_S) <- c("sex", "dilution")

# add columns to df 
df.convert_full <- cbind(df.convert,df.convert_S)

# add in block information 
# this was not reproducable to count but I don't know how else to do it
# first 118 rows are block A, second 127 rows are block B, the third 138 rows are block C, and the last 133 rows are block D
df.convert_full$Block <- rep(c("A","B", "C", "D"), c(118, 127, 138, 129))


```

Plot survival curve with only females

```{r}

df.convert_fem <- df.convert_full[which(df.convert_full$sex != "male"),]

# order treatments 
df.convert_fem <- df.convert_fem %>% 
  mutate(dilution = factor(dilution, levels = c("CCM", "0.01", "0.1", "1", "3", "6")))

# change to not have confidence intervals in this one so you can see them 
df_fit_fem<- survfit(Surv(dead, status) ~ dilution, data=df.convert_fem)
ggsurvplot(df_fit_fem, size = 5,
          pval = FALSE, conf.int = FALSE,
          legend = "bottom",
          font.tickslab = c(14),
          font.x = c(16),
          font.y = c(16),
          font.t = c(16),
          ggtheme = theme_light(),
          title = expression(paste("Female",italic(" D. innubila "), "Injected with Dilutions of Passage 4 DiNV")),
          legend.title="Treatment",
          legend.labs=c("CCM", "0.01 FFU", "0.1 FFU", "1 FFU", "3 FFU", "6 FFU"),
          font.legend = c(14),
          palette = c("#E7E1EF", "#C994C7","#DF65B0","#E7298A" ,"#980043", "#67001F")) + ylab("Survival Proporation") + xlab("Days post injection")
        
          

```

Plot survival curve with only males

```{r}

df.convert_m <- df.convert_full[which(df.convert_full$sex != "female"),]

# order treatments 
df.convert_m <- df.convert_m %>% 
  mutate(dilution = factor(dilution, levels = c("CCM", "0.01", "0.1", "1", "3", "6")))

# change to not have confidence intervals in this one so you can see them 
df_fit_m<- survfit(Surv(dead, status) ~ dilution, data=df.convert_m)
ggsurvplot(df_fit_m, size = 5,
          pval = FALSE, conf.int = FALSE,
          legend = "bottom",
          font.tickslab = c(14),
          font.x = c(16),
          font.y = c(16),
          font.t = c(16),
          ggtheme = theme_light(),
          title = expression(paste("Male",italic(" D. innubila "), "Injected with Dilutions of Passage 4 DiNV")),
          legend.title="Treatment",
          legend.labs=c("CCM", "0.01 FFU", "0.1 FFU", "1 FFU", "3 FFU", "6 FFU"),
          font.legend = c(14),
          palette = c("#E7E1EF", "#C994C7","#DF65B0","#E7298A" ,"#980043", "#67001F")) + ylab("Survival Proporation") + xlab("Days post injection")
        
          
```

Look at median survival time

```{r}
# females 
surv_median(df_fit_fem, combine = FALSE)
# males 
surv_median(df_fit_m, combine = FALSE)
```

Look at model for males comparing dilutions

```{r}
male_1 <- coxph(Surv(dead, status) ~ dilution + Block, data=df.convert_m)
summary(male_1)
```

Look at model for females comparing dilutions

```{r}
female_1 <- coxph(Surv(dead, status) ~ dilution + Block, data=df.convert_fem)
summary(female_1)
```

Compare all dilutions with sex as an interaction term Using continuous variables

```{r}
# make dilution as a numeric/continuous variable
# make a duplicate column 
df.convert_full$numeric.dilution <- df.convert_full$dilution
# set CCM to a dose of 0
df.convert_full$numeric.dilution[which(df.convert_full$numeric.dilution == "CCM")] = 0
# make all values numeric 
df.convert_full$numeric.dilution <- as.numeric(df.convert_full$numeric.dilution)

df_all_fit<- coxph(Surv(dead, status) ~ sex*numeric.dilution + Block, data=df.convert_full)
summary(df_all_fit)
```

Compare above model with no sex

```{r}
df_all_nosex_fit<- coxph(Surv(dead, status) ~ numeric.dilution + Block, data=df.convert_full)
summary(df_all_nosex_fit)

```

Model with numeric dilution, block, and sex with no interaction

```{r}
df_all_noint_fit<- coxph(Surv(dead, status) ~ numeric.dilution + Block + sex, data=df.convert_full)
summary(df_all_noint_fit)
```

Compare all three models together

```{r}

models <- list(df_all_fit, df_all_nosex_fit, df_all_noint_fit)

model.names <- c( 'model with sex and dilution interaction', 'model without sex as a variable', 'model with sex but no interaction')

aictab(cand.set = models, modnames = model.names)

# best model is df_all_noint_fit but not by much 
summary(df_all_noint_fit)
```

Compare 0.1 FFU, 1 FFU, 3 FFU, and 6 FFU together

```{r}
# remove CCM or 0 dose treatment 
df_highdils <- df.convert_full[which(df.convert_full$numeric.dilution != 0),]
# remove 0.01 FFU dose treatment 
df_highdils <- df_highdils[which(df_highdils$numeric.dilution != 0.01),]

# dilution and sex interaction
df_highdils_fit <- coxph(Surv(dead, status) ~ numeric.dilution*sex + Block, data=df_highdils)
summary(df_highdils_fit)
# dilution and sex no interaction 
df_highdils_fitni <- coxph(Surv(dead, status) ~ numeric.dilution + sex + Block, data=df_highdils)
summary(df_highdils_fitni)
# dilution no sex
df_highdils_fitns <- coxph(Surv(dead, status) ~ numeric.dilution + Block, data=df_highdils)
summary(df_highdils_fitns)





models2 <- list(df_highdils_fit, df_highdils_fitni, df_highdils_fitns )

model.names2 <- c( 'dilution and sex interaction', 'dilution and sex no interaction', 'just dilution')

aictab(cand.set = models2, modnames = model.names2)

# best model is df_highdils_fitni no interaction 
summary(df_highdils_fitni)
```

Compare 0.01 FFU and CCM together

```{r}
# remove 1, 3, 6, 0.1 dose treatment 
df_lowdils <- df.convert_full[which(df.convert_full$numeric.dilution != 0.1),]
# remove 0.01 FFU dose treatment 
df_lowdils <- df_lowdils[which(df_lowdils$numeric.dilution != 1),]
df_lowdils <- df_lowdils[which(df_lowdils$numeric.dilution != 3),]
df_lowdils <- df_lowdils[which(df_lowdils$numeric.dilution != 6),]

# dilution and sex interaction 
df_lowdils_fit<- coxph(Surv(dead, status) ~ numeric.dilution*sex + Block, data=df_lowdils)
summary(df_lowdils_fit)
# dilution and sex no interaction 
df_lowdils_fitni <- coxph(Surv(dead, status) ~ numeric.dilution + sex + Block, data=df_lowdils)
summary(df_lowdils_fitni)
# dilution no sex
df_lowdils_fitns <- coxph(Surv(dead, status) ~ numeric.dilution + Block, data=df_lowdils)
summary(df_lowdils_fitns)


models3 <- list(df_lowdils_fit, df_lowdils_fitni, df_lowdils_fitns )

model.names3 <- c( 'dilution and sex interaction', 'dilution and sex no interaction', 'just dilution')

aictab(cand.set = models3, modnames = model.names3)

# best model does not have sex df_lowdils_fitns
summary(df_lowdils_fitns)
```


Changing color pallet and lines for paper 

Female 

```{r}
df_fit_fem<- survfit(Surv(dead, status) ~ dilution, data=df.convert_fem)
ggsurvplot(df_fit_fem, size = 2.5,
          pval = FALSE, conf.int = FALSE,
          legend = "bottom",
          font.tickslab = c(14),
          font.x = c(16),
          font.y = c(16),
          font.t = c(16),
          ggtheme = theme_light(),
          title = expression(paste("Female",italic(" D. innubila "), "Injected with Dilutions of Passage 4 DiNV")),
          legend.title="Treatment",
          legend.labs=c("CCM", "0.01 FFU", "0.1 FFU", "1 FFU", "3 FFU", "6 FFU"),
          font.legend = c(14),
          palette = c("#009E73", "#648FFF","#785EF0","#DC267F" ,"#FE6100", "#FFB000")) + ylab("Survival Proporation") + xlab("Days post injection")
        
```

Male

```{r}
df_fit_m<- survfit(Surv(dead, status) ~ dilution, data=df.convert_m)
ggsurvplot(df_fit_m, size = 2.5,
          pval = FALSE, conf.int = FALSE,
          legend = "bottom",
          font.tickslab = c(14),
          font.x = c(16),
          font.y = c(16),
          font.t = c(16),
          ggtheme = theme_light(),
          title = expression(paste("Male",italic(" D. innubila "), "Injected with Dilutions of Passage 4 DiNV")),
          legend.title="Treatment",
          legend.labs=c("CCM", "0.01 FFU", "0.1 FFU", "1 FFU", "3 FFU", "6 FFU"),
          font.legend = c(14),
          palette = c("#009E73", "#648FFF","#785EF0","#DC267F" ,"#FE6100", "#FFB000")) + ylab("Survival Proporation") + xlab("Days post injection")
```


