---
title: "p4-dilutions-time"
format: gfm
editor: visual
---

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(Rmisc)
library(ggpubr)
library(rstatix)
library(AICcmodavg)
library(ggtext)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width=8, fig.height = 5)
```

Load in dataset

```{r}
Cq_values <- read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/qPCR_analysis/20240228-viral-load-dilutions/viral-dil-time-sheet.csv")
```

Separate male and female for now

```{r}
# males
Cq_values_M <- Cq_values[which(Cq_values$sex == "male"),]
# females
Cq_values_F <- Cq_values[which(Cq_values$sex == "female"),]
```

**Female Analysis First**

Look at raw Cq values

```{r}
ggplot(Cq_values_F, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(~dilution) 
```

TPI looks remarkably consistant, PIF3 is all over the place but this isn't separated out by day which would make differences.

Look at raw Cq values and also separate by day

```{r}
ggplot(Cq_values_F, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(dilution ~ day) 
```

This is showing more patterns that we sort of expect, but we have to do the delta Cq comparison to the TPI primer to be sure

Calculating female delta Cq

```{r}
# need to organize by name of the sample I think
Cq_values_F <- Cq_values_F[order(Cq_values_F$sample_ID),]
# this worked to order by the sample number 
# and the PIF3 value is first in the order and TPI second

# Separate that dataframe, incriminating by 2, every number between 1-192 (number of rows in dataframe)
Cq_values_F$Cq[seq(1,192,2)] # these are the PIF 3 Cq #s
Cq_values_F$Cq[seq(2,192,2)] # these are the TPI primer Cq #s 

# make the delta Cq by subtracting the PIF 3 values from the TPI primer values
# and this is saved as a vector in R 
delta_Cqs_F <- Cq_values_F$Cq[seq(2,192,2)] - Cq_values_F$Cq[seq(1,192,2)]
#vector
delta_Cqs_F

# Make a new dataframe that only has one row per sample by getting rid of the rows with the TPI primer
Cq_values_F_Delta <- Cq_values_F[which(Cq_values_F$primer == "PIF3"),]

# And then add in the delta Cqs as a new column
Cq_values_F_Delta$delta_Cq <- delta_Cqs_F

# add a column with 2^ delta Cq
Cq_values_F_Delta$delta_Cq_2 <- 2^(delta_Cqs_F)

```

Plot 2\^delta Cq

```{r}
ledgend_title = "Virus Delivery"
ggplot(Cq_values_F_Delta, aes(y= delta_Cq_2, x=day)) +
  geom_boxplot(outlier.shape=NA, aes(fill=dilution))  + 
  theme_light() + scale_fill_manual(ledgend_title, values=c("#E7E1EF", "#C994C7", "#CE1256")) + 
  geom_dotplot(aes(fill=dilution), binaxis='y', stackdir='center', dotsize=0.6, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=6), legend.title=element_text(size=7)) +
  scale_x_discrete(labels=c("day0" = "0 days", "day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) + 
  labs(title = "Comparing Viral Load in Female _D. innubila_ Across Early Infection",y = "Relative amount of DiNV genome to host genome", x = "Days Post Injection") +
  theme(legend.position = c(0.93, 0.15), legend.background = element_rect(linetype="solid", colour ="black"),plot.title = element_markdown())
```

**Male Analysis Second**

Look at raw Cq values and also separate by day

```{r}
ggplot(Cq_values_M, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(dilution ~ day) 
```

Looks pretty similar to the female results

Calculating male delta Cq

```{r}
# need to organize by name of the sample I think
Cq_values_M <- Cq_values_M[order(Cq_values_M$sample_ID),]
# this worked to order by the sample number 
# and the PIF3 value is first in the order and TPI second

# notice that there is one NA in the data that I forgot to set as 40 for the Cq 

Cq_values_M[is.na(Cq_values_M)] <- 40

# Separate that dataframe, incriminating by 2, every number between 1-192 (number of rows in dataframe)
Cq_values_M$Cq[seq(1,192,2)] # these are the PIF 3 Cq #s
Cq_values_M$Cq[seq(2,192,2)] # these are the TPI primer Cq #s 

# make the delta Cq by subtracting the PIF 3 values from the TPI primer values
# and this is saved as a vector in R 
delta_Cqs_M <- Cq_values_M$Cq[seq(2,192,2)] - Cq_values_M$Cq[seq(1,192,2)]
#vector
delta_Cqs_M

# Make a new dataframe that only has one row per sample by getting rid of the rows with the TPI primer
Cq_values_M_Delta <- Cq_values_M[which(Cq_values_M$primer == "PIF3"),]

# And then add in the delta Cqs as a new column
Cq_values_M_Delta$delta_Cq <- delta_Cqs_M

# add a column with 2^ delta Cq
Cq_values_M_Delta$delta_Cq_2 <- 2^(delta_Cqs_M)

```

Plot 2\^delta Cq

```{r}
ggplot(Cq_values_M_Delta, aes(y= delta_Cq_2, x=day)) +
  geom_boxplot(outlier.shape=NA, aes(fill=dilution))  + 
  theme_light() + scale_fill_manual(ledgend_title, values=c("#E7E1EF", "#C994C7", "#CE1256")) + 
  geom_dotplot(aes(fill=dilution), binaxis='y', stackdir='center', dotsize=0.6, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=6), legend.title=element_text(size=7)) +
  scale_x_discrete(labels=c("day0" = "0 days", "day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) + 
  labs(title = expression("Comparing Viral Load in Male _D. innubila_ Across Early Infection"),y = "Relative amount of DiNV genome to host genome", x = "Days Post Injection") +
  theme(legend.position = "none",
        plot.title = element_markdown())
```

**Delta Delta Analysis**

Females

```{r}
# this might be a little mess to do 
# separate out the day 0 samples to start with 
F_0 <- Cq_values_F_Delta[which(Cq_values_F_Delta$day == "day0"),]
# separate out dilutions?
F_0_3 <- F_0[which(F_0$dilution == "3 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(F_0_3$delta_Cq)
# -9.25875
F_0_01 <- F_0[which(F_0$dilution == "0.1 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(F_0_01$delta_Cq)
# -13.38625
F_0_001 <- F_0[which(F_0$dilution == "0.01 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(F_0_001$delta_Cq)
# -15.4525

# main data sheet without day 0 
F <- Cq_values_F_Delta[which(Cq_values_F_Delta$day != "day0"),]
# separate out just dilution 3 
F_3 <- F[which(F$dilution == "3 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_3 <- F_3$delta_Cq - -9.25875
delta_delta_3

# add as column 
F_3$delta_delta_Cq <- delta_delta_3

# do 2^ delta delta 
F_3$delta_delta_Cq_2 <- 2^(delta_delta_3)

# separate out just dilution 0.1 
F_01 <- F[which(F$dilution == "0.1 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_01 <- F_01$delta_Cq - -13.38625
delta_delta_01

# add as column 
F_01$delta_delta_Cq <- delta_delta_01

# do 2^ delta delta 
F_01$delta_delta_Cq_2 <- 2^(delta_delta_01)

# separate out just dilution 0.01 
F_001 <- F[which(F$dilution == "0.01 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_001 <- F_001$delta_Cq - -15.4525
delta_delta_001

# add as column 
F_001$delta_delta_Cq <- delta_delta_001

# do 2^ delta delta 
F_001$delta_delta_Cq_2 <- 2^(delta_delta_001)

# add all of the days back together 
Female_delta_delta <- rbind(F_001, F_01, F_3)

```

Doing linear models for statistical tests Female use delta delta Cq

```{r}
# make a dilution column that is numeric
Female_delta_delta$numeric.dilution <- as.numeric(gsub(" FFU", "", Female_delta_delta$dilution))
# make a day column that is numeric
Female_delta_delta$numeric.day <- as.numeric(gsub("day", "", Female_delta_delta$day))

# linear model 
# do stats on just the delta delta Cq, not the 2^ transformed 
female_full_model <- lm(delta_delta_Cq~numeric.dilution*numeric.day, data = Female_delta_delta)
summary(aov(female_full_model))
# no interaction 
female_full_modelni <- lm(delta_delta_Cq~numeric.dilution+numeric.day, data = Female_delta_delta)
summary(aov(female_full_modelni))

modelsF <- list(female_full_model, female_full_modelni)

model.namesF <- c( ' dilution day interaction', 'dilution day')

aictab(cand.set = modelsF, modnames = model.namesF)
# no interaction model female_delta_full_modelni is the best 
summary(female_full_modelni)

# look at just 0.01 FFU dilution 
female_001_model <- lm(delta_delta_Cq~numeric.day, data = subset(Female_delta_delta, numeric.dilution == "0.01"))
summary(aov(female_001_model))
```

Plot box plots

```{r}
legend_title <- "Virus Delivery"
ggplot(Female_delta_delta, aes(y= delta_delta_Cq_2, x=day)) + geom_boxplot(outlier.shape=NA,aes(fill=dilution)) +  
  scale_fill_manual(legend_title, values=c("#E7E1EF", "#C994C7", "#CE1256")) + 
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.6, position=position_dodge(0.8),aes(fill=dilution)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=10), legend.title=element_text(size=11)) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  labs(title = "Comparing DiNV Increase in Female Flies \nInjected with Various Titers Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection") +
  theme(legend.position = c(0.9, 0.15), legend.background = element_rect(linetype="solid", colour ="black")) 
```

Males Calculate delta delta Cq

```{r}
# this might be a little mess to do 
# separate out the day 0 samples to start with 
M_0 <- Cq_values_M_Delta[which(Cq_values_M_Delta$day == "day0"),]
# separate out dilutions?
M_0_3 <- M_0[which(M_0$dilution == "3 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(M_0_3$delta_Cq)
# -9.4875
M_0_01 <- M_0[which(M_0$dilution == "0.1 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(M_0_01$delta_Cq)
# -12.63125
M_0_001 <- M_0[which(M_0$dilution == "0.01 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(M_0_001$delta_Cq)
# -13.40625

# main data sheet without day 0 
M <- Cq_values_M_Delta[which(Cq_values_M_Delta$day != "day0"),]
# separate out just dilution 3 
M_3 <- M[which(M$dilution == "3 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_3_M <- M_3$delta_Cq - -9.4875
delta_delta_3_M

# add as column 
M_3$delta_delta_Cq <- delta_delta_3_M

# do 2^ delta delta 
M_3$delta_delta_Cq_2 <- 2^(delta_delta_3_M)

# separate out just dilution 0.1 
M_01 <- M[which(M$dilution == "0.1 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_01_M <- M_01$delta_Cq - -12.63125
delta_delta_01_M

# add as column 
M_01$delta_delta_Cq <- delta_delta_01_M

# do 2^ delta delta 
M_01$delta_delta_Cq_2 <- 2^(delta_delta_01_M)

# separate out just dilution 0.01 
M_001 <- M[which(M$dilution == "0.01 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_001_M <- M_001$delta_Cq - -13.40625
delta_delta_001_M

# add as column 
M_001$delta_delta_Cq <- delta_delta_001_M

# do 2^ delta delta 
M_001$delta_delta_Cq_2 <- 2^(delta_delta_001_M)

# add all of the days back together 
Male_delta_delta <- rbind(M_001, M_01, M_3)
```

Doing linear models for statistical tests Male

```{r}
# make a dilution column that is numeric
Male_delta_delta$numeric.dilution <- as.numeric(gsub(" FFU", "", Male_delta_delta$dilution))
# make a day column that is numeric
Male_delta_delta$numeric.day <- as.numeric(gsub("day", "", Male_delta_delta$day))

# linear model 
# do stats on just the delta delta Cq, not the 2^ transformed 
male_full_model <- lm(delta_delta_Cq~numeric.dilution*numeric.day, data = Male_delta_delta)
summary(aov(male_full_model))
# no interaction 
male_full_modelni <- lm(delta_delta_Cq~numeric.dilution+numeric.day, data = Male_delta_delta)
summary(aov(male_full_modelni))

modelsM <- list(male_full_model, male_full_modelni)

model.namesM <- c( ' dilution day interaction', 'dilution day')

aictab(cand.set = modelsM, modnames = model.namesM)
# no interaction model is the best
summary(male_full_modelni)

# look at just 0.01 FFU dilution 
Male_001_model <- lm(delta_delta_Cq~numeric.day, data = subset(Male_delta_delta, numeric.dilution == "0.01"))
summary(aov(Male_001_model))
```

Plot box plots

```{r}
legend_title <- "Virus Delivery"
ggplot(Male_delta_delta, aes(y= delta_delta_Cq_2, x=day)) + geom_boxplot(outlier.shape=NA,aes(fill=dilution)) +  
  scale_fill_manual(legend_title, values=c("#E7E1EF", "#C994C7", "#CE1256")) + 
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.6, position=position_dodge(0.8),aes(fill=dilution)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=10), legend.title=element_text(size=11)) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  labs(title = "Comparing DiNV Increase in Male Flies \nInjected with Various Titers Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection") +
  theme(legend.position = "none")
```

Combine males and females and look at a linear model considering all for the delta delta Cq

```{r}
all_data <- rbind(Female_delta_delta, Male_delta_delta)

# model with interaction between day, dilution and sex
full_model_int <- lm(delta_delta_Cq~numeric.dilution*numeric.day*sex, data = all_data)
summary(aov(full_model_int))

# model with interaction between day and dilution, but not sex
full_model_nsint <- lm(delta_delta_Cq~numeric.dilution*numeric.day + sex, data = all_data)
summary(aov(full_model_nsint))

# model with interaction between day and dilution, and model not including sex 
full_model_ns <- lm(delta_delta_Cq~numeric.dilution*numeric.day, data = all_data)
summary(aov(full_model_ns))


# try special package for comparing AIC 
models <- list( full_model_ns, full_model_nsint, full_model_int)

model.names <- c( 'model without sex', 'model without sex as interaction', 'full model')

aictab(cand.set = models, modnames = model.names)

# model without sex is the best fit 
```

Look at male and female for each dilution by day

3 FFU dilution

```{r}
# male
M_3_dil <- Male_delta_delta[which(Male_delta_delta$dilution == "3 FFU"),]
# female
F_3_dil <- Female_delta_delta[which(Female_delta_delta$dilution == "3 FFU"),]

FFU_3_dil <- rbind(M_3_dil, F_3_dil)


legend_title = "Sex"
#Plot 
ggplot(FFU_3_dil, aes(y= delta_delta_Cq_2, x=day, fill=sex)) + 
  geom_boxplot()  + 
  scale_fill_manual(legend_title, values=c("#E7298A" ,  "#67001F")) +
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=11), legend.title=element_text(size=12)) +
  theme(legend.position = c(0.9, 0.2), legend.background = element_rect(linetype="solid", colour ="black")) +
  labs(title = "Comparing DiNV Increase in Male and Female Flies \nInjected with 3 FFU DiNV Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection")

```

0.1 FFU dilution

```{r}
# male
M_01_dil <- Male_delta_delta[which(Male_delta_delta$dilution == "0.1 FFU"),]
# female
F_01_dil <- Female_delta_delta[which(Female_delta_delta$dilution == "0.1 FFU"),]

FFU_01_dil <- rbind(M_01_dil, F_01_dil)


legend_title = "Sex"
#Plot 
ggplot(FFU_01_dil, aes(y= delta_delta_Cq_2, x=day, fill=sex)) + 
  geom_boxplot()  + 
  scale_fill_manual(legend_title, values=c("#E7298A" ,  "#67001F")) +
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=11), legend.title=element_text(size=12)) +
  theme(legend.position = "none") +
  labs(title = "Comparing DiNV Increase in Male and Female Flies \nInjected with 0.1 FFU DiNV Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection")


```

0.01 FFU dilution

```{r}
# male
M_001_dil <- Male_delta_delta[which(Male_delta_delta$dilution == "0.01 FFU"),]
# female
F_001_dil <- Female_delta_delta[which(Female_delta_delta$dilution == "0.01 FFU"),]

FFU_001_dil <- rbind(M_001_dil, F_001_dil)


legend_title = "Sex"
#Plot 
ggplot(FFU_001_dil, aes(y= delta_delta_Cq_2, x=day, fill=sex)) + 
  geom_boxplot()  + 
  scale_fill_manual(legend_title, values=c("#E7298A" ,  "#67001F")) +
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=11), legend.title=element_text(size=12)) +
  theme(legend.position = "none") +
  labs(title = "Comparing DiNV Increase in Male and Female Flies \nInjected with 0.01 FFU DiNV Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection")
```

**Adding in days 6, 7, and 9 for males** Not sure if I will include this, I don't feel comfortable combining two experiments like this and there were larger differences in the day 0 PIF 3 Cqs than I would have expected

*this was not included in final analysis because we really can't add in extra samples that had a different day 0*

Load in dataset

```{r}
Cq_values_extra <- read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/qPCR_analysis/20240228-viral-load-dilutions/20240313-extra-0.01.csv")
```

Look at raw Cq values

```{r}
ggplot(Cq_values_extra, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(~day) 
```

These look pretty similar to above, I wonder if all the flies aren't getting infected...

Calculating extra delta Cq and Plotting

```{r}
# need to organize by name of the sample I think
Cq_values_extra <- Cq_values_extra[order(Cq_values_extra$sample_ID),]
# this worked to order by the sample number 
# and the TPI value is first in the order and PIF 3 second

# Separate that dataframe, incriminating by 2, every number between 1-48 (number of rows in dataframe)
Cq_values_extra$Cq[seq(1,48,2)] # these are the TPI Cq #s
Cq_values_extra$Cq[seq(2,48,2)] # these are the PIF 3 primer Cq #s 

# make the delta Cq by subtracting the PIF 3 values from the TPI primer values
# and this is saved as a vector in R 
delta_Cqs_extra <- Cq_values_extra$Cq[seq(1,48,2)] - Cq_values_extra$Cq[seq(2,48,2)]
#vector
delta_Cqs_extra

# Make a new dataframe that only has one row per sample by getting rid of the rows with the TPI primer
Cq_values_extra_Delta <- Cq_values_extra[which(Cq_values_extra$primer == "PIF 3"),]

# And then add in the delta Cqs as a new column
Cq_values_extra_Delta$delta_Cq <- delta_Cqs_extra

# add a column with 2^ delta Cq
Cq_values_extra_Delta$delta_Cq_2 <- 2^(delta_Cqs_extra)


# plot and use a log 10 scale for the y axis 
ggplot(Cq_values_extra_Delta, aes(y= delta_Cq_2, x=day)) + geom_boxplot()  + theme_linedraw() + geom_point() + scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x)))

```

Delta delta analysis

```{r}
# this might be a little mess to do 
# separate out the day 0 samples to start with 
E_0 <- Cq_values_extra_Delta[which(Cq_values_extra_Delta$day == "day0"),]

# mean of day 0 delta Cq
mean(E_0$delta_Cq)
# -10.274

# main data sheet without day 0 
E <- Cq_values_extra_Delta[which(Cq_values_extra_Delta$day != "day0"),]

# subtract average day 0 delta Cq from the delta Cq of all others 
delta_delta_extra <- E$delta_Cq - -10.274
delta_delta_extra

# add as column 
E$delta_delta_Cq <- delta_delta_extra

# do 2^ delta delta 
E$delta_delta_Cq_2 <- 2^(delta_delta_extra)


# plot
legend_title <- "Virus Delivery"

ggplot(E, aes(y= delta_delta_Cq_2, x=day)) + geom_boxplot() +  
  theme_light() + geom_point() +
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=14)) +
  scale_x_discrete(labels=c("day6" = "6 days", "day7" = "7 days", "day9" = "9 days")) +
  labs(title = "Comparing Viral Titer in Male Flies \nInjected with Various Titers Over Early Infection",y = "2^delta delta Cq", x = "Days Since Injection")
```

Combine this data with the male data from the other experiment

```{r}
# male data
#head(M_001_dil)
# to combine datasets I'll need all the same columns 
# don't need plate column, sex column 
#M_001_dil_s <- M_001_dil[,c(1,3:4,6:13)]

# combine extra and original data 
# add all of the days back together 
#All_Male_delta_delta <- rbind(M_001_dil_s, E)

# plot
#legend_title <- "Virus Delivery"

#ggplot(All_Male_delta_delta, aes(y= delta_delta_Cq_2, x=day, fill=dilution)) + geom_boxplot() +  
  #scale_fill_manual(legend_title, values=c( "#67001F")) + 
  #theme_light() + 
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + 
  #scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  #theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=14)) +
  #scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days", "day6" = "6 days", "day7" = "7 days", "day9" = "9 days")) +
  #labs(title = "Comparing Viral Titer in Male Flies \nInjected with Various Titers Over Infection",y = "2^delta delta Cq", x = "Days Since Injection")
```


Female delta delta modified for paper figures 

changing 

```{r}
legend_title <- "Virus Delivery"
ggplot(Female_delta_delta, aes(y= delta_delta_Cq_2, x=day)) + geom_boxplot(outlier.shape=NA,aes(fill=dilution)) +  
  scale_fill_manual(legend_title, values=c("#648FFF", "#DC267F", "#FFB000")) + 
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.6, position=position_dodge(0.8),aes(fill=dilution)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=10), legend.title=element_text(size=11)) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  labs(title = "Comparing DiNV Increase in Female Flies \nInjected with Various Titers Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection") +
  theme(legend.position = c(0.9, 0.15), legend.background = element_rect(linetype="solid", colour ="black")) 
```


Males delta delta color change 

```{r}
legend_title <- "Virus Delivery"
ggplot(Male_delta_delta, aes(y= delta_delta_Cq_2, x=day)) + geom_boxplot(outlier.shape=NA,aes(fill=dilution)) +  
  scale_fill_manual(legend_title, values=c("#648FFF", "#DC267F", "#FFB000")) + 
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.6, position=position_dodge(0.8),aes(fill=dilution)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=10), legend.title=element_text(size=11)) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  labs(title = "Comparing DiNV Increase in Male Flies \nInjected with Various Titers Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection") +
  theme(legend.position = "none")
```


Female one delta 

```{r}
ledgend_title = "Virus Delivery"
ggplot(Cq_values_F_Delta, aes(y= delta_Cq_2, x=day)) +
  geom_boxplot(outlier.shape=NA, aes(fill=dilution))  + 
  theme_light() + scale_fill_manual(ledgend_title, values=c("#648FFF", "#DC267F", "#FFB000")) + 
  geom_dotplot(aes(fill=dilution), binaxis='y', stackdir='center', dotsize=0.6, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=6), legend.title=element_text(size=7)) +
  scale_x_discrete(labels=c("day0" = "0 days", "day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) + 
  labs(title = "Comparing Viral Load in Female _D. innubila_ Across Early Infection",y = "Relative amount of DiNV genome to host genome", x = "Days Post Injection") +
  theme(legend.position = c(0.93, 0.15), legend.background = element_rect(linetype="solid", colour ="black"),plot.title = element_markdown())
```


Male one delta 

```{r}
ggplot(Cq_values_M_Delta, aes(y= delta_Cq_2, x=day)) +
  geom_boxplot(outlier.shape=NA, aes(fill=dilution))  + 
  theme_light() + scale_fill_manual(ledgend_title, values=c("#648FFF", "#DC267F", "#FFB000")) + 
  geom_dotplot(aes(fill=dilution), binaxis='y', stackdir='center', dotsize=0.6, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=6), legend.title=element_text(size=7)) +
  scale_x_discrete(labels=c("day0" = "0 days", "day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) + 
  labs(title = expression("Comparing Viral Load in Male _D. innubila_ Across Early Infection"),y = "Relative amount of DiNV genome to host genome", x = "Days Post Injection") +
  theme(legend.position = "none",
        plot.title = element_markdown())
```



