---
title: "20240330-Dinn-Dv-1-qPCR"
format: gfm
editor: visual
---

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(Rmisc)
library(rstatix)
library(AICcmodavg)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width=7, fig.height = 6)
```

Load in dataset

```{r}
# lambda and PIF 3
Cq_values <- read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/qPCR_analysis/20240330-Dinn-Dv-1/20240330-Dinn-Dv-1.csv")
# load in RPL11 and TPI data
Cq_2 <- read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/qPCR_analysis/20240330-Dinn-Dv-1/RPL11-TPI.csv")
```

Look at raw Cq values

```{r}
ggplot(Cq_values, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(~cell_type ~day ~infection) 
```

Controls look very similar and infections look very similar for each cell line which is good I think.

```{r}
# use the variance function, and calculates the variance in Cq by the unique.name (each sample/primer has 3 Cq values to 
# calculate the variance by)
Cq_values$Cq_var <- ave(Cq_values$Cq, Cq_values$unique_name, FUN=var)

# use the mean function, and calculates the mean of Cq values by the unique.name (each sample/primer has 3 Cq values to 
# calculate the mean by)
Cq_values$Cq_mean <- ave(Cq_values$Cq, Cq_values$unique_name, FUN=mean)

# Keep all rows where the replicate is 1 (or you could do 2 or 3)
# make into new Df so we keep the original with all the Cq values
Cq_values_1rep <- Cq_values[which(Cq_values$replicate == "1"),]

# histogram of all variances
ggplot(Cq_values_1rep, aes(x=Cq_var)) + geom_histogram(bins = 50)
# Ok some variation is really big, but most is fine 
# I do think this would be the controls for PIF 3 because some can be 40 and some can be like 35 and mess up the variance 
# If there is 1 Cq value that is vastly different than the other 2 technical replicatates I am going to remove it 

# first order df by the Cq variance 
ordered <- Cq_values[order(Cq_values$Cq_var, decreasing = TRUE),]
head(ordered)
# change rownames
rownames(ordered) <- NULL
# what rows have a Cq value that is very different than the other 2 replicates 
# 2, 4, 8, 11, 15, 18, 19, 24, 33, 45, 46, 49
#remove those rows with negative indexing 
ordered_2 <- ordered[-c(2, 4, 8, 11, 15, 18, 19, 24, 33, 45, 46, 49), ]
# now to re-do the variance and the means 
ordered_2$Cq_var <- ave(ordered_2$Cq, ordered_2$unique_name, FUN=var)
ordered_2$Cq_mean <- ave(ordered_2$Cq, ordered_2$unique_name, FUN=mean)
# this seems to have worked good 
# now I need to keep only 1 replicate row 
# remove "duplicated" rows based on the unique_name column: each technical replicate should have the same unique_name
Cq_values_var_1rep <- ordered_2[!duplicated(ordered_2$unique_name), ]
# re-look at histogram of variances
# there will probably be a few high ones because for some samples there was not an obvious 1 Cq value different than the others 
ggplot(Cq_values_var_1rep, aes(x=Cq_var)) + geom_histogram(bins = 50)
# this I think is a lot better and what I will move forward with 
```

Calculating Delta Cqs between Lambda and PIF3

```{r}
# need to order the dataset 
# try ordering by primer first then by sample 
Cq_values_var_1rep <- Cq_values_var_1rep[order(Cq_values_var_1rep$primer),]
Cq_values_var_1rep
# then order by sample 
Cq_values_var_1rep <- Cq_values_var_1rep[order(Cq_values_var_1rep$sample),]
Cq_values_var_1rep

# this consistently LAMBDA first, then PIF 3 
nrow(Cq_values_var_1rep)

# Separate that dataframe, incriminating by 2, every number between 1-64 (number of rows in dataframe)
Cq_values_var_1rep$Cq_mean[seq(1,64,2)] # these are the LAMBDA Cq means 
Cq_values_var_1rep$Cq_mean[seq(2,64,2)] # these are the PIF 3 primer Cq means 

# make the delta Cq by subtracting the PIF 3 values from the LAMBDA primer values
# and this is saved as a vector in R 
delta_Cqs <- Cq_values_var_1rep$Cq_mean[seq(1,64,2)] - Cq_values_var_1rep$Cq_mean[seq(2,64,2)]
#vector
delta_Cqs

# Keep only rows that are PIF3 because now we have two rows per sample
Cq_values1rep_Delta <- Cq_values_var_1rep[which(Cq_values_var_1rep$primer == "PIF 3"),]
# And then add in the delta Cqs as a new column
Cq_values1rep_Delta$delta_Cq <- delta_Cqs

# do 2^ delta Cq
Cq_values1rep_Delta$delta_Cq_2 <- 2^(delta_Cqs)

# plot 
ggplot(Cq_values1rep_Delta, aes(y= delta_Cq_2, x=cell_type, fill=infection)) + geom_boxplot()  + theme_linedraw()  + facet_grid(~day)  + scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x)))


```

Calculating delta delta Cq

Controls

```{r}
# separate out control and DiNV samples 
control <- Cq_values1rep_Delta[which(Cq_values1rep_Delta$infection == "control"),]
DiNV <- Cq_values1rep_Delta[which(Cq_values1rep_Delta$infection != "control"),]
 
# separate out the day 0 samples to start with 
day_0_cc <- control[which(control$day == "day0"),]
# separate out cell type Dv-1 
day_0_cc_Dv1 <- day_0_cc[which(day_0_cc$cell_type == "DV-1"),]
# find mean of Delta Cq (not 2^delta) for Dv-1
mean(day_0_cc_Dv1$delta_Cq)
# -10.77958
# separate out cell type Myd88 
day_0_cc_Dinn <- day_0_cc[which(day_0_cc$cell_type != "DV-1"),]
# find mean of Delta Cq (not 2^delta) for Dinn
mean(day_0_cc_Dinn$delta_Cq)
# -9.9275

# separate out day 5 only from controls 
day_5_cc <- control[which(control$day != "day0"),]
# separate out Dv-1
day_5_cc_Dv1 <- day_5_cc[which(day_5_cc$cell_type == "DV-1"),]

# subtract average day 0 delta Cq for Dv-1 from the delta Cq of all others 
day_5_cc_Dv1_delta_delta <- day_5_cc_Dv1$delta_Cq - -10.77958
day_5_cc_Dv1_delta_delta

# add as column 
day_5_cc_Dv1$delta_delta_Cq <- day_5_cc_Dv1_delta_delta

# do 2^ delta delta 
day_5_cc_Dv1$delta_delta_Cq_2 <- 2^(day_5_cc_Dv1_delta_delta)

# separate out Dinn
day_5_cc_Dinn <- day_5_cc[which(day_5_cc$cell_type != "DV-1"),]

# subtract average day 0 delta Cq for Dinn from the delta Cq of all others 
day_5_cc_Dinn_delta_delta <- day_5_cc_Dinn$delta_Cq - -9.9275
day_5_cc_Dinn_delta_delta

# add as column 
day_5_cc_Dinn$delta_delta_Cq <- day_5_cc_Dinn_delta_delta

# do 2^ delta delta 
day_5_cc_Dinn$delta_delta_Cq_2 <- 2^(day_5_cc_Dinn_delta_delta)

# combine controls back together  
Control_delta_delta <- rbind(day_5_cc_Dv1, day_5_cc_Dinn)
```

Delta Delta calculations for DiNV samples

```{r}
 
# separate out the day 0 samples to start with 
day_0_DiNV <- DiNV[which(DiNV$day == "day0"),]
# separate out cell type Dv-1 
day_0_DiNV_Dv1 <- day_0_DiNV[which(day_0_DiNV$cell_type == "DV-1"),]
# find mean of Delta Cq (not 2^delta) for Dv-1
mean(day_0_DiNV_Dv1$delta_Cq)
# 3.690833
# separate out cell type Dinn 
day_0_DiNV_Dinn <- day_0_DiNV[which(day_0_DiNV$cell_type != "DV-1"),]
# find mean of Delta Cq (not 2^delta) for Dv-1
mean(day_0_DiNV_Dinn$delta_Cq)
# 3.756667

# separate out day 5 only from DiNVs
day_5_DiNV <- DiNV[which(DiNV$day != "day0"),]
# separate out Dv-1
day_5_DiNV_Dv1 <- day_5_DiNV[which(day_5_DiNV$cell_type == "DV-1"),]

# subtract average day 0 delta Cq for Dv-1 from the delta Cq of all others 
day_5_DiNV_Dv1_delta_delta <- day_5_DiNV_Dv1$delta_Cq -  3.690833
day_5_DiNV_Dv1_delta_delta

# add as column 
day_5_DiNV_Dv1$delta_delta_Cq <- day_5_DiNV_Dv1_delta_delta

# do 2^ delta delta 
day_5_DiNV_Dv1$delta_delta_Cq_2 <- 2^(day_5_DiNV_Dv1_delta_delta)

# separate out Dinn
day_5_DiNV_Dinn <- day_5_DiNV[which(day_5_DiNV$cell_type != "DV-1"),]

# subtract average day 0 delta Cq for Dinn from the delta Cq of all others 
day_5_DiNV_Dinn_delta_delta <- day_5_DiNV_Dinn$delta_Cq - 3.756667
day_5_DiNV_Dinn_delta_delta

# add as column 
day_5_DiNV_Dinn$delta_delta_Cq <- day_5_DiNV_Dinn_delta_delta

# do 2^ delta delta 
day_5_DiNV_Dinn$delta_delta_Cq_2 <- 2^(day_5_DiNV_Dinn_delta_delta)

# combine controls back together  
DiNV_delta_delta <- rbind(day_5_DiNV_Dv1, day_5_DiNV_Dinn)
```

Combine control and DiNV back together and plot delta delta Cq

```{r}
delta_delta <- rbind(Control_delta_delta, DiNV_delta_delta)

legend_title <- "Infection"

# plot 
ggplot(delta_delta, aes(y= delta_delta_Cq_2, x=cell_type, fill=infection)) + geom_boxplot(outlier.shape=NA)  + theme_light()  + scale_fill_manual(legend_title, values=c("#C994C7", "#CE1256"), labels= c("Cell culture medium", "DiNV")) + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + scale_x_discrete(labels=c("DV-1" = "Dv-1", "Dinn" = "Dinn-1"))+ 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=14), legend.position = "bottom") +
  labs(title = "Comparing Increase in Virus in Dv-1 \nand Dinn-1 Cells 5 Days Post Innoculation",y = "Relative DiNV Genome Increase", x = "Cell Line")
```

Pairwise t test with lambda primers

```{r}
stat.test <- delta_delta %>%
  group_by(infection) %>%
  t_test(delta_delta_Cq_2 ~ cell_type) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
stat.test
```

Linear model instead of t-test (better)

```{r}
# using lambda as control
# compare model with infection and cell type
# use only the delta delta Cq, not 2^
Dv_Dinn_l_model <- lm(delta_delta_Cq~infection*cell_type, data = delta_delta)
summary(aov(Dv_Dinn_l_model))

# model without interaction 
Dv_Dinn_l_modelni <- lm(delta_delta_Cq~infection+cell_type, data = delta_delta)
summary(aov(Dv_Dinn_l_modelni))

# model without cell type 
Dv_Dinn_l_modelnc <- lm(delta_delta_Cq~infection, data = delta_delta)
summary(aov(Dv_Dinn_l_modelnc))


# try special package for comparing AIC 
modelsl <- list( Dv_Dinn_l_model, Dv_Dinn_l_modelni, Dv_Dinn_l_modelnc)

model.namesl <- c( 'interaction', 'no interaction', 'no cell type')

aictab(cand.set = modelsl, modnames = model.namesl)

# best model is no cell type 
summary(aov(Dv_Dinn_l_modelnc))

```

Now compare Pif 3 Cqs with TPI/RPL11 for delta delta Cq

Look at variance in control primer Cqs

```{r}

# use the variance function, and calculates the variance in Cq by the unique.name (each sample/primer has 3 Cq values to 
# calculate the variance by)
Cq_2$Cq_var <- ave(Cq_2$Cq, Cq_2$unique_name, FUN=var)

# use the mean function, and calculates the mean of Cq values by the unique.name (each sample/primer has 3 Cq values to 
# calculate the mean by)
Cq_2$Cq_mean <- ave(Cq_2$Cq, Cq_2$unique_name, FUN=mean)

# Keep all rows where the replicate is 1 (or you could do 2 or 3)
# make into new Df so we keep the original with all the Cq values
Cq_2_1rep <- Cq_2[which(Cq_2$replicate == "1"),]

# histogram of all variances
ggplot(Cq_2_1rep, aes(x=Cq_var)) + geom_histogram(bins = 50)

# all of these are very low, great
```

Delta Cq with Cell primers

```{r}
# separate out just PIF 3 samples from Cq_values
Cq_pif <- Cq_values_var_1rep[which(Cq_values_var_1rep$primer == "PIF 3"),]
# combine datasets 

cell_Cqs <- rbind(Cq_pif, Cq_2_1rep)

# need to order the dataset 
# try ordering by primer first then by sample 
cell_Cqs <- cell_Cqs[order(cell_Cqs$primer),]
cell_Cqs
# then order by sample 
cell_Cqs <- cell_Cqs[order(cell_Cqs$sample),]
cell_Cqs

# this consistently PIF 3 first, then control primers 
nrow(cell_Cqs)

# Separate that dataframe, incriminating by 2, every number between 1-64 (number of rows in dataframe)
cell_Cqs$Cq_mean[seq(1,64,2)] # these are the PIF 3 Cq means 
cell_Cqs$Cq_mean[seq(2,64,2)] # these are the cell primer Cq means 

# make the delta Cq by subtracting the PIF 3 values from the cell primer values
# and this is saved as a vector in R 
delta_cell_Cqs <- cell_Cqs$Cq_mean[seq(2,64,2)] - cell_Cqs$Cq_mean[seq(1,64,2)]
#vector
delta_cell_Cqs

# Keep only rows that are PIF3 because now we have two rows per sample
cell_Cqs_Delta <- cell_Cqs[which(cell_Cqs$primer == "PIF 3"),]
# And then add in the delta Cqs as a new column
cell_Cqs_Delta$delta_Cq <- delta_cell_Cqs

# do 2^ delta Cq
cell_Cqs_Delta$delta_Cq_2 <- 2^(delta_cell_Cqs)

# plot 
ggplot(cell_Cqs_Delta, aes(y= delta_Cq_2, x=cell_type, fill=infection)) + geom_boxplot()  + theme_linedraw()  + facet_grid(~day)  + scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x)))


```

Very similar to what we saw with Lambda, which should be expected!

Calculating delta delta Cq with cell primers

Controls

```{r}
# separate out control and DiNV samples 
Ccontrol <- cell_Cqs_Delta[which(cell_Cqs_Delta$infection == "control"),]
CDiNV <- cell_Cqs_Delta[which(cell_Cqs_Delta$infection != "control"),]
 
# separate out the day 0 samples to start with 
Cday_0_cc <- Ccontrol[which(Ccontrol$day == "day0"),]
# separate out cell type Dv-1 
Cday_0_cc_Dv1 <- Cday_0_cc[which(Cday_0_cc$cell_type == "DV-1"),]
# find mean of Delta Cq (not 2^delta) for Dv-1
mean(Cday_0_cc_Dv1$delta_Cq)
# -9.065417
# separate out cell type Dinn 
Cday_0_cc_Dinn <- Cday_0_cc[which(Cday_0_cc$cell_type != "DV-1"),]
# find mean of Delta Cq (not 2^delta) for Dinn
mean(Cday_0_cc_Dinn$delta_Cq)
# -11.57

# separate out day 5 only from controls 
Cday_5_cc <- Ccontrol[which(Ccontrol$day != "day0"),]
# separate out Dv-1
Cday_5_cc_Dv1 <- Cday_5_cc[which(Cday_5_cc$cell_type == "DV-1"),]

# subtract average day 0 delta Cq for Dv-1 from the delta Cq of all others 
Cday_5_cc_Dv1_delta_delta <- Cday_5_cc_Dv1$delta_Cq - -9.065417
Cday_5_cc_Dv1_delta_delta

# add as column 
Cday_5_cc_Dv1$delta_delta_Cq <- Cday_5_cc_Dv1_delta_delta

# do 2^ delta delta 
Cday_5_cc_Dv1$delta_delta_Cq_2 <- 2^(Cday_5_cc_Dv1_delta_delta)

# separate out Dinn
Cday_5_cc_Dinn <- Cday_5_cc[which(Cday_5_cc$cell_type != "DV-1"),]

# subtract average day 0 delta Cq for Dinn from the delta Cq of all others 
Cday_5_cc_Dinn_delta_delta <- Cday_5_cc_Dinn$delta_Cq - -11.57
Cday_5_cc_Dinn_delta_delta

# add as column 
Cday_5_cc_Dinn$delta_delta_Cq <- Cday_5_cc_Dinn_delta_delta

# do 2^ delta delta 
Cday_5_cc_Dinn$delta_delta_Cq_2 <- 2^(Cday_5_cc_Dinn_delta_delta)

# combine controls back together  
CControl_delta_delta <- rbind(Cday_5_cc_Dv1, Cday_5_cc_Dinn)
```

Delta Delta calculations for DiNV samples

```{r}
# separate out the day 0 samples to start with 
Cday_0_DiNV <- CDiNV[which(CDiNV$day == "day0"),]
# separate out cell type Dv-1 
Cday_0_DiNV_Dv1 <- Cday_0_DiNV[which(Cday_0_DiNV$cell_type == "DV-1"),]
# find mean of Delta Cq (not 2^delta) for Dv-1
mean(Cday_0_DiNV_Dv1$delta_Cq)
# 3.305833
# separate out cell type Dinn 
Cday_0_DiNV_Dinn <- Cday_0_DiNV[which(Cday_0_DiNV$cell_type != "DV-1"),]
# find mean of Delta Cq (not 2^delta) for Dinn
mean(Cday_0_DiNV_Dinn$delta_Cq)
# 1.433333

# separate out day 5 only from DiNVs
Cday_5_DiNV <- CDiNV[which(CDiNV$day != "day0"),]
# separate out Dv-1
Cday_5_DiNV_Dv1 <- Cday_5_DiNV[which(Cday_5_DiNV$cell_type == "DV-1"),]

# subtract average day 0 delta Cq for Dv-1 from the delta Cq of all others 
Cday_5_DiNV_Dv1_delta_delta <- Cday_5_DiNV_Dv1$delta_Cq -  3.305833
Cday_5_DiNV_Dv1_delta_delta

# add as column 
Cday_5_DiNV_Dv1$delta_delta_Cq <- Cday_5_DiNV_Dv1_delta_delta

# do 2^ delta delta 
Cday_5_DiNV_Dv1$delta_delta_Cq_2 <- 2^(Cday_5_DiNV_Dv1_delta_delta)

# separate out Dinn
Cday_5_DiNV_Dinn <- Cday_5_DiNV[which(Cday_5_DiNV$cell_type != "DV-1"),]

# subtract average day 0 delta Cq for Dinn from the delta Cq of all others 
Cday_5_DiNV_Dinn_delta_delta <- Cday_5_DiNV_Dinn$delta_Cq - 1.433333
Cday_5_DiNV_Dinn_delta_delta

# add as column 
Cday_5_DiNV_Dinn$delta_delta_Cq <- Cday_5_DiNV_Dinn_delta_delta

# do 2^ delta delta 
Cday_5_DiNV_Dinn$delta_delta_Cq_2 <- 2^(Cday_5_DiNV_Dinn_delta_delta)

# combine controls back together  
CDiNV_delta_delta <- rbind(Cday_5_DiNV_Dv1, Cday_5_DiNV_Dinn)
```

Combine control and DiNV back together and plot delta delta Cq

```{r}
Cdelta_delta <- rbind(CControl_delta_delta, CDiNV_delta_delta)

legend_title <- "Infection"

# plot 
ggplot(Cdelta_delta, aes(y= delta_delta_Cq_2, x=cell_type, fill=infection)) + geom_boxplot(outlier.shape=NA)  + theme_light()  + scale_fill_manual(legend_title, values=c("#C994C7", "#CE1256"), labels= c("Cell culture medium", "DiNV")) + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + scale_x_discrete(labels=c("DV-1" = "Dv-1", "Dinn" = "Dinn-1"))+ 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=14), legend.position = "bottom") +
  labs(title = "Comparing Increase in Virus in Dv-1 \nand Dinn-1 Cells 5 Days Post Innoculation",y = "Relative DiNV Genome Increase", x = "Cell Line")
```

find mean increase for Dv-1 samples

```{r}
# separate out Dv-1
Dv <- Cdelta_delta[which(Cdelta_delta$cell_type == "DV-1"),]
# separate out DiNV infected 
DvD <- Dv[which(Dv$infection == "DiNV"),]
meanD <- mean(DvD$delta_delta_Cq_2)
meanD
```

find mean increase for Dinn samples

```{r}
# separate out Dinn
Di <- Cdelta_delta[which(Cdelta_delta$cell_type == "Dinn"),]
# separate out DiNV infected 
DiD <- Di[which(Di$infection == "DiNV"),]
meanDi <- mean(DiD$delta_delta_Cq_2)
meanDi
```

Pairwise t test with cell primers

```{r}
stat.test.c <- Cdelta_delta %>%
  group_by(infection) %>%
  t_test(delta_delta_Cq_2 ~ cell_type) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
stat.test.c
```

Linear model instead of t-test (better)

```{r}
# using cell primers as control
# compare model with infection and cell type
# use only the delta delta Cq, not 2^
Dv_Dinn_c_model <- lm(delta_delta_Cq~infection*cell_type, data = Cdelta_delta)
summary(aov(Dv_Dinn_c_model))
# model without interaction 
Dv_Dinn_c_modelni <- lm(delta_delta_Cq~infection+cell_type, data = Cdelta_delta)
summary(aov(Dv_Dinn_c_modelni))
# model without cell type 
Dv_Dinn_c_modelnc <- lm(delta_delta_Cq~infection, data = Cdelta_delta)
summary(aov(Dv_Dinn_c_modelnc))


# try special package for comparing AIC 
modelsc <- list( Dv_Dinn_c_model, Dv_Dinn_c_modelni, Dv_Dinn_c_modelnc)

model.namesc <- c( 'interaction', 'no interaction', 'no cell type')

aictab(cand.set = modelsc, modnames = model.namesc)

# interaction model is the best 
summary(aov(Dv_Dinn_c_model))
```


Making figures with different color pallets for paper (colorblind friendly)

Lambda comparison 

changing colors 

no legend position because in Myd88 figure 

```{r}
legend_title <- "Infection"

# plot 
ggplot(delta_delta, aes(y= delta_delta_Cq_2, x=cell_type, fill=infection)) + geom_boxplot(outlier.shape=NA)  + theme_light()  + scale_fill_manual(legend_title, values=c("#009E73", "#CC79A7"), labels= c("Cell culture medium", "DiNV")) + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + scale_x_discrete(labels=c("DV-1" = "Dv-1", "Dinn" = "Dinn-1"))+ 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=14), legend.position = "none") +
  labs(title = "Comparing Increase in Virus in Dv-1 \nand Dinn-1 Cells 5 Days Post Innoculation",y = "Relative DiNV Genome Increase", x = "Cell Line")
```


Cell primers 
Putting legend in the plot and changing colors 

```{r}
legend_title <- "Infection"

# plot 
ggplot(Cdelta_delta, aes(y= delta_delta_Cq_2, x=cell_type, fill=infection)) + geom_boxplot(outlier.shape=NA)  + theme_light()  + scale_fill_manual(legend_title, values=c("#009E73", "#CC79A7"), labels= c("Cell culture medium", "DiNV")) + geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + scale_x_discrete(labels=c("DV-1" = "Dv-1", "Dinn" = "Dinn-1"))+ 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=14), legend.position = c(0.2, 0.88), 
        legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
  labs(title = "Comparing Increase in Virus in Dv-1 \nand Dinn-1 Cells 5 Days Post Innoculation",y = "Relative DiNV Genome Increase", x = "Cell Line")
```









